%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape    
    We have seen that computer programming is an art, \\ 
    because it applies accumulated knowledge to the world, \\ 
    because it requires skill and ingenuity, and especially \\
    because it produces objects of beauty.} \\ \medskip
    --- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
\end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgments}

My thoughts go to my family: \\
\indent thank you \emph{big} dad Alessandro, \\
\indent thank you \emph{loving} mom Grazia,\\
\indent thank you  \emph{lively} sister Angela, \\
without your love, warmth, support, encouragement 
and patience \emph{I would not be who now I am}.
\\\\
Thank you Niccol\`o for the great \emph{being} you are.
\\\\
Thank you Filippo for our long cycling trips and our lasting
friendship, although the physical distance among us.
\\\\
Thank you Francesco, Arash, Michele and Samuele for your help,
your teachings and how unique your \flqq be my friend\frqq\,has been.
\\\\
Thank you master Riccardo for your enthusiasm in the \emph{Judo} you teach.
\\\\
Thank you professors Donatella and Giovanni for your kind and motivating help,
words and passion for what you do.

\vfill

\noindent Just \emph{You}, nothing else.
%\noindent\emph{You}, grappled to a \emph{Pepsi cork}, which is my \emph{heart}.

\vfill


\iffalse
Put your acknowledgments here.

Many thanks to everybody who already sent me a postcard!

Regarding the typography and other help, many thanks go to Marco 
Kuhlmann, Philipp Lehman, Lothar Schlesier, Jim Young, Lorenzo 
Pantieri and Enrico Gregorio\footnote{Members of GuIT (Gruppo 
Italiano Utilizzatori di \TeX\ e \LaTeX )}, J\"org Sommer, 
Joachim K\"ostler, Daniel Gottschlag, Denis Aydin, Paride 
Legovini, Steffen Prochnow, Nicolas Repp, Hinrich Harms, 
 Roland Winkler, J\"org Weber, 
 and the whole \LaTeX-community for support, ideas and 
 some great software.

\bigskip

\noindent\emph{Regarding \mLyX}: The \mLyX\ port was intially done by 
\emph{Nicholas Mariette} in March 2009 and continued by 
\emph{Ivo Pletikosi\'c} in 2011. Thank you very much for your 
work and the contributions to the original style.
\fi

\endgroup

\vfill

\begin{flushright}{\slshape    
    \vskip-3cm 
    \rightskip-3cm
    E ritorn\`o dalla volpe.\\
    -- Addio -- disse.\\
    -- Addio -- disse la volpe.\\
    -- Ecco il mio segreto. \`E molto semplice:\\
        \flqq non si vede bene che con il cuore,
        l'essenziale \`e invisibile agli occhi\frqq\\
    -- L'essenziale \`e invisibile agli occhi -- ripet\`e il Piccolo Principe, per ricordarselo.\\
    -- \`E il tempo che tu hai perduto per la tua rosa che ha fatto la tua rosa cos\`i importante.\\
    -- \`E il tempo che ho perduto per la mia rosa\ldots -- sussurr\`o il Piccolo Principe per ricordarselo.\\
    E, riverso sull'erba, pianse.\\
    \quad
    } 
\end{flushright}

